# Citations
The R package 'citations' comprises functions that can be used to analyze citations cases, i.e. citations within documents that cite a particular (or several) studies.

The functions contained should help to 

1. Identify the citing document and collect meta information: `identify_citing_document()`
1. Extract text from those PDFs and save in text files: `extract_text()`
2. Identify whether texts really cite the particular study of interest: `identify_study_in_text()`
3. Delete certain sections in texts to facilitate citation case extraction: `delete_reference_section()` and `delete_running_heads()`
5. Clean texts to facilitate citation case extraction: `clean_text()`
8. Extract citations cases from these documents: `extract_citation_cases()`
9. Analyze the extracted citation cases: to be continued...


# How to...

## Prerequesites

* You need to install xpdf into the folder C:\Program Files\xpdf

## First steps
First install/load the package.
```{r, echo=T, eval=F}
install.packages("devtools")
devtools::install_github("paulbauer/citations")
library(citations)
```

Second, set the working directory to the folder that, in turn, contains the folders (e.g. "Beck 1995", "Tumasjan 2010") that contain the PDFs.
```{r, echo=T, eval=F}
setwd("C:/Users/paul/Google Drive/Research/2016_Quality_of_citations/data")
```




## Example: A single study (20 documents)
```{r, echo=T, eval=F}
identify_citing_document(folder = "Beck 1995", number = 10)
extract_text(folder = "Beck 1995", number = 20)
identify_study(study.title = "What to Do (and Not to Do) with Time-Series Cross-Section Data",
                 folder = "Beck 1995"#
                 ,  number = 20)
delete_reference_section(folder = "Beck 1995", number = 20)
delete_running_heads(folder = "Beck 1995", number = 20)
clean_text(folder = "Beck 1995", number = 20)
extract_citation_cases(folder = "Beck 1995",
                         authorname = "Beck, Katz",
                         studyyear = "1995",
                         number = 20) # scope = 2
```


## Example: A single study
```{r, echo=T, eval=F}
identify_citing_document(folder = "Beck 1995")
extract_text(folder = "Beck 1995")
identify_study(study.title = "What to Do (and Not to Do) with Time-Series Cross-Section Data",
                 folder = "Beck 1995"#
                 )
delete_reference_section(folder = "Beck 1995")
delete_running_heads(folder = "Beck 1995")
clean_text(folder = "Beck 1995")
extract_citation_cases(folder = "Beck 1995",
                         authorname = "Beck, Katz",
                         studyyear = "1995"
                         ) # scope = 2
```


## Example: Several studies
* Info on studies should be saved in a  table
* Then we can loop over the table content

```{r, echo=T, eval=F}

publicationdata <- read.table("publicationdata.txt", sep=";",
                                header = T, stringsAsFactors = F)

  for (i in 1:nrow(publicationdata)){

    folder <- publicationdata[i,1]
    study.title <- publicationdata[i,2]
    authorname <- publicationdata[i,3]
    studyyear <- publicationdata[i,4]

    # extract_text(folder = folder)
    # identify_study(study.title = study.title, folder = folder)
    # delete_reference_section(folder = folder)
    # clean_text(folder = folder)
    extract_citation_cases(folder = folder, authorname = authorname, studyyear = studyyear)

  }
```
