# Citations
The R package 'citations' comprises functions that can be used to analyze citations cases, i.e. citations within documents that cite a particular (or several) studies.

The functions contained should help to 

1. Extract text from those PDFs and save in text files: `extract_text()`
2. Identify whether texts really cite the particular study of interest: `identify_study_in_text()`
3. Delete certain sections in texts to facilitate citation case extraction: `delete_reference_section()` and `delete_running_heads()`
5. Clean texts to facilitate citation case extraction: `clean_text()`
8. Extract citations cases from these documents: `extract_citation_cases()`
9. Analyze the extracted citation cases: to be continued...


# How to...

## Prerequesites

* You need to install xpdf into the folder C:\Program Files\xpdf

## Install and loading the package

```{r, echo=T, eval=F}
install.packages("devtools")
devtools::install_github("paulbauer/citations")
library(citations)
```



## Set working directory (that contains folder with the PDF documents)

```{r, echo=T, eval=F}
setwd("C:\\Users\\paul\\Desktop\\citationdata")
```


## Example: Single study

# SINGLE STUDY (20 documents)
```{r, echo=T, eval=F}
extract_text(folder = "Beck 1995", number = 20)
identify_study(study.title = "What to Do (and Not to Do) with Time-Series Cross-Section Data",
                 folder = "Beck 1995"#
                 ,  number = 20)
delete_reference_section(folder = "Beck 1995", number = 20)
delete_running_heads(folder = "Beck 1995", number = 20)
clean_text(folder = "Beck 1995", number = 20)
extract_citation_cases(folder = "Beck 1995",
                         authorname = "Beck, Katz",
                         studyyear = "1995",
                         number = 20) # scope = 2
```


# SINGLE STUDY (ALL documents)
```{r, echo=T, eval=F}
extract_text(folder = "Beck 1995")
identify_study(study.title = "What to Do (and Not to Do) with Time-Series Cross-Section Data",
                 folder = "Beck 1995"#
                 )
delete_reference_section(folder = "Beck 1995")
delete_running_heads(folder = "Beck 1995")
clean_text(folder = "Beck 1995")
extract_citation_cases(folder = "Beck 1995",
                         authorname = "Beck, Katz",
                         studyyear = "1995"
                         ) # scope = 2


  extract_text(folder = "Ermisch 2009")
  identify_study(study.title = "What to Do (and Not to Do) with Time-Series Cross-Section Data",
                 folder = "Ermisch 2009"#
  )
  delete_reference_section(folder = "Ermisch 2009")
  delete_running_heads(folder = "Ermisch 2009")
  clean_text(folder = "Ermisch 2009")
  extract_citation_cases(folder = "Ermisch 2009",
                         authorname = "Ermisch, Gambetta, Laurie, Siedler, Uhrig",
                         studyyear = "2009"
  ) # scope = 2

```


# SEVERAL STUDIES
* Info on studies should be saved in a  table
* Then we can loop over the table content
```{r, echo=T, eval=F}

publicationdata <- read.table("publicationdata.txt", sep=";",
                                header = T, stringsAsFactors = F)

  for (i in 1:nrow(publicationdata)){

    folder <- publicationdata[i,1]
    study.title <- publicationdata[i,2]
    authorname <- publicationdata[i,3]
    studyyear <- publicationdata[i,4]

    # extract_text(folder = folder)
    # identify_study(study.title = study.title, folder = folder)
    # delete_reference_section(folder = folder)
    # clean_text(folder = folder)
    extract_citation_cases(folder = folder, authorname = authorname, studyyear = studyyear)

  }
```
